Shaping the Perception of a Robot as a Human-Like Agent by Cognitive Priming

### 2015-11-03 main.tex
Re-organized and mostly rewritten intro

Robotics researchers often tend to consider that anthropomorphism only describes
a static set of human-like features of a robot (like its shape, its speech
capabilities, facial expressions, etc.).
Following~\citet{fink_anthropomorphism_2012}, we refer to these characteristics
as the \emph{anthropomorphic design} of the robot, and we call
\emph{anthropomorphism} the general \emph{social phenomenon} that emerges from
the (real or imagined) interaction between a non-human agent (a robot in our
case) and a human~\citep{persson_anthropomorphism_2000}.  According
to~\citet{epley_when_2008}, anthropomorphism includes the human perception of
emotional states, motivations, intentions in the non-human agent, and tends to
ascribe those qualities to it. As such, the dynamic and socio-cognitive
dimensions of anthropomorphism are essential to understand the complex bonds
that human users build with artificial agents such as
robots~\citep{lemaignan2014dynamics}. Therefore anthropomorphism appears to be a
useful \emph{proxy} to study these intricate psychological
bonds. Going in that direction, this article introduces a new way to assess
anthropomorphic attributions and provides in particular new insights on
the impact of \emph{cognitive priming} on anthropomorphic
attributions.

\subsection{Role of Context in Human-Robot Interaction}

The study of context in human-robot interaction (HRI) is not new.
\textit{Context of use} is defined by the ISO as \textit{``users, tasks,
equipment (hardware, software and materials), and the physical and social
environments in which a product is used''}.  Accordingly, there are various
aspects that can impact how a robot is experienced. The real or imagined
\textit{purpose} of a robot, including the application context in which it is
typically used (\eg office environment, home, rehabilitation center, public
space, \etc) as well as the task context (\eg serious or playful task;
composition of human-robot team; number of users) and the (social) role in which
the robot is used / experienced are to be taken into account.

Notably, when the context of use of a robot is social, entertaining or playful,
it will enhance anthropomorphism compared to when the context is a routine or
focused serious task (security, rescue, \etc).  \citet{joosse_what_2013} showed,
for instance, that when the same robot ({\sc nao}) was used in a different task
context (cleaning task \vs tour guide), users ascribed different personalities
to the robot. \citet{goetz_cooperation_2002} revealed a link between a robot's
context of use and people's perception of the robot: the authors found that
people prefer a serious robot for serious tasks and a less serious robot for
more playful tasks. \citet{kaplan_free_2000} discussed the role of uselessness
in the design of robots. The author argues that artificial pets like AIBO have
no real purpose, in a sense that they do not provide any kind of service or, in
other words, \textit{``they are not doing what you tell them to do''} and
\textit{``they might refuse the order of its owner''}. It may be this very
aspect that increases people's tendency to anthropomorphize the robot. According
to Kaplan, it is in our daily use language that we tend to attribute intentions
to devices that are not doing their job well.

These previous studies investigated users' perception of robots in
prototypical situations (routine \vs playful, typically boring \vs typically
intellectual) and their results nicely supports our general intuition.
In this article, we show that much more subtle variations in the task context do
lead to very noticeable effects as well.

\subsection*{Measuring Anthropomorphism}

Measuring anthropomorphism raises
interesting methodological challenges, if only because people's perception of
robots change over the duration of their interaction with the
robots~\citep{lemaignan2014dynamics}. \citet{takayama_perspectives_2012}
discusses the effects of these changes and differentiates between what she calls
an \emph{in-the-moment} and a \emph{reflective} perspective on agency: an
\emph{in-the-moment} perspective would refer to one's most immediate
response/sense in a given situation whereas a \emph{reflective} perspective would
indicate one's reaction/sense of a situation based on a more in-depth
consideration and contemplation.  In the first phase of interaction with a
robot, people might respond ``mindlessly'' instead of responding
consciously~\citep{nass_machines_2000}, and only after what is called the
\emph{familiarization} period, one might respond in a more reflective manner
instead of an in-the-moment reaction: \citet{reeves_media_1996} showed for
instance that participants interacting with technologies in ways close to how they
would behave towards people, would still deny any human-like interaction in a post-hoc
questionnaire, illustrating this reflective perspective.

Yet, most of the experiments that investigate anthropomorphism in HRI are based
on post-interaction, closed questionnaires or rating scales.  The Godspeed
questionnaire~\citep{bartneck_measurement_2008}, and in particular, its subpart
focused on anthropomorphism, is the main validated questionnaire to assess
anthropomorphism. On 5-points semantic differential scales, people are asked to
rate the following constructs: fake \vs natural, machinelike \vs humanlike,
unconscious \vs conscious, artificial \vs lifelike, moving rigidly \vs moving
elegantly. Because the concept of ``human-likeness'' itself is complex and
abstract, \citet{kahn_jr._robotic_2006} suggest to ask for more concrete
constructs that are typical or unique of the concept of ``human-likeness'', and
\citet{ruijten_introducing_2014} propose for instance a 25-item questionnaire to
measure various concrete aspects of human-likeness. Other
researchers~\citep{zlotowski2014dimensions,salem2015would} also applied a
two-dimensional scale measuring \emph{Human Nature} and \emph{Uniquely Human}
traits of robots, based on an original proposal by \citet{haslam2008attributing}.

We explore here a different approach to measure anthropomorphism, based on
eye-tracking.  By looking at the eyes fixation on the head of the robot, we have
been able to demonstrate significantly different patterns in two experimental
conditions subtly different in terms of human-likeness. This eye-tracking based
metric is behavioural and hence does not suffer post-hoc reconstruction by the
participants, possibly providing a better, ``noise-less'', access to the human
attitude towards the machine, and, as such, a better proxy to the cognitive and
affective bonds that the human establishes with the robot.

@@

This leads to a second-order effect that we also evidence here: depending on the
cognitive priming, the tendency to attribute human-like characteristics to robot
evolves in significantly different ways. After observing a robot in a
context that pre-suppose deeper cognitive capabilities, people increase
their general tendency to anthropomorphize robots compared to a shallow
cognitive priming, even though the robot appearance and visual behaviour is the
exact same.

>>>>

Researchers in robotics often tend to see anthropomorphism as the static set of
human-like features of a robot (like its shape, its speech capabilities, facial
expressions, etc.). Following~\citet{fink_anthropomorphism_2012}, we refer to
these characteristics as the \emph{anthropomorphic design} of the robot, and we
call instead \emph{anthropomorphism} the \emph{social phenomenon} that arises
from the (real or imagined) interaction between a non-human agent (a robot in
our case) and a human~\citep{persson_anthropomorphism_2000}. As such,
anthropomorphism includes the perception of emotional states, motivations,
intentions in the non-human agent, and, conversely, the ascriptions of those
qualities to them~\citep{epley_when_2008}.

Being such a central ingredient of the psychological ``sauce'' that binds humans to
robots, the field of human-robot interaction (HRI) has long~\fixme{ref?}
attempted to manipulate the anthropomorphic attributes we ascribe to robots.
While modifying the anthropomorphic \emph{design} of robots is conceptually
easy, influencing the underlying psychological phenomenon is less
obvious, although likely more effective.

One way consists in manipulating the context of use of the robot: when social,
entertaining or playful, it elicits more anthropomorphic attributions compared
routine or focused, serious situations (security, rescue, \etc).
\citet{joosse_what_2013} showed, for instance, that when the same robot (a {\sc
nao} humanoid) was used in a different task context (cleaning task \vs tour
guide), users would ascribe different personalities to the robot. Along the same
line, \citet{goetz_cooperation_2002} found that people prefer a serious robot
for serious tasks and a less serious robot for more playful tasks. Also related,
\citet{kaplan_free_2000} discussed how the uselessness of the robot impacts our
tendency to anthropomorphize it: artificial pets like AIBO have no real purpose,
in a sense that they do not provide any kind of service, and this very aspect
may increase people's tendency to anthropomorphize them.

We may hypothesize that these results are a consequence of different
expectations we form about the robot depending on the context: when a machine is
embedded in a social situation, for instance, we expect it to exhibit certain
social skills. However, the experiments conducted so far explore prototypical
opposite situations (routine \vs playful, typically boring \vs typically
intellectual): while they do confirm that the interaction context impacts our
perception of the robot, the stark contrast of the conditions makes it difficult
to finely analyse the psychological determinants that explain the different
perception of the robot.\fixme{references are needed} In particular, existing
techniques to measure anthropomorphism traditionally rely on questionnaires that
may suffer post-hoc reconstruction (we come back on this hereafter).

This article attempts to address this issue with a different experimental design
and a novel methodology to measure \emph{in-the-moment} anthropomorphic attributions.
Our approach is built on a combination of cognitive priming and eye-tracking,
and evidences specific gaze fixations on the head of the robot when the
participants are primed with higher expectations regarding the cognitive skills
of the robot.

\subsection{Cognitive Priming and Interaction Situation}

\textbf{TBD: Kshitij} at the end of this section, the reader must be convinced
that:
\begin{itemize}
    \item cognitive priming is a well established technique, grounded in the
        litterature,
    \item our terminology ('shallow' vs 'deep') is meaningful, and grounded
    \item our cognitive priming would impact the perception of the interaction
        situation as machine-like vs human-like (to support our manipulation
        check!)
\end{itemize}

\subsection{Measuring Anthropomorphism}

Assessing anthropomorphic attributions during a human-robot interaction is an
indirect mean to explore the intricate -- and dynamic -- bonds that build up
between the robot and its human partner. As such, a lot of literature in HRI has
investigated means of measuring anthropomorphism. This indeed raises interesting
methodological challenges, if only because of high inter-personal variations in
one's tendency to anthropomorphize~\fixme{ref?}, and the dynamic nature of
anthropomorphic attributions~\citep{lemaignan2014dynamics}.

One phenomenon requires particular attention: well established in the
psychological literature in terms of \emph{impulsive} or \emph{automatic} \vs
\emph{reflective} behaviours~\citep{strack2004reflective}, the way we behave and
perceive \emph{in-the-moment} is significantly different from what results from
a thoughtful (and slower) reflection on the situation. As noted by
\citet{takayama_perspectives_2012} for instance, this also impacts our
perspective on agency, and in fact, as shown by~\citet{nass_machines_2000}, in
the first phase of interaction with a robot, people might respond mindlessly,
and only after the so-called \emph{familiarization} period, one might respond in
a more conscious, reflective manner. The typical illustration of this mechanism
is found in~\citet{reeves_media_1996} who showed that participants who would
interact with technologies in ways similar to how they interact with people,
would however {\it a posteriori} typically deny having performed an human-like
interaction.

Yet, most of the methodologies available to assess anthropomorphic ascriptions
in HRI are limited to post-interaction measurement, mainly in the form of closed
questionnaires or rating scales. The Godspeed
questionnaire~\citep{bartneck_measurement_2008}, and in particular, its subpart
focused on anthropomorphism, is the main validated questionnaire to assess
anthropomorphism. On 5-points semantic differential scales, people are asked to
rate the following constructs: fake \vs natural, machinelike \vs humanlike,
unconscious \vs conscious, artificial \vs lifelike, moving rigidly \vs moving
elegantly. Because the concept of ``human-likeness'' itself is complex and
abstract, \citet{kahn_jr._robotic_2006} suggest to ask for more concrete
constructs that are typical or unique of the concept of ``human-likeness'', and
\citet{ruijten_introducing_2014} propose for instance a 25-item questionnaire to
measure various concrete aspects of human-likeness. Other
researchers~\citep{zlotowski2014dimensions,salem2015would} also applied a
two-dimensional scale measuring \emph{Human Nature} and \emph{Uniquely Human}
traits of robots, based on an original proposal by
\citet{haslam2008attributing}.

We explore hereafter a different approach to measure anthropomorphism, based on
eye-tracking. By looking at the eyes fixation on the head of the robot, we show
significantly different gazing behaviours in two experimental conditions that
differ in terms of the (projected) robot's human-likeness. This eye-tracking
based metric is behavioural -- hence not prone to post-hoc reconstruction by the
participants -- and \emph{in-the-moment} -- mitigating the impact of a reflective
elaboration on the nature of the human-robot interaction. We believe that these
two improvements over existing techniques provide a solid ground for the study
of the socio-cognitive bonds that establish between humans and robots when they
interact.

@@

This leads to a second-order effect that we also evidence here: depending on the
cognitive priming, the tendency to attribute human-like characteristics to robot
evolves in significantly different ways. After observing a robot in a
context that presupposes deeper cognitive capabilities, people increase
their general tendency to anthropomorphize robots compared to a shallow
cognitive priming, even though the robot appearance and visual behaviour is the
exact same.
